Light and growth form interact to shape stomatal ratio among British angiosperms

### 2017-10-05 ms/ms.Rnw
updated results

\subsection*{Interactions between light and Raunki\ae r life form determine stomatal ratio}

Overall, $\mathrm{SR_{even}}$ increased with \el, but there was a significant interaction ($\Delta\mathrm{AIC} > 2$, Table~\ref{table:srAIC}) between Raunki\ae r life form and \el~(Fig.~\ref{fig:SRmultReg}). Both life form and \el~significantly increased model fit, though \el~had a markedly larger effect on model AIC (Table~\ref{table:srAIC}). The significant interaction is caused by different slopes between life forms. Among life forms with the overall greatest \el~(therophytes, hemicryptophytes, and chamaephytes, see Fig.~\ref{fig:lf-light}), there was a strong positive relationship between \el~and $\mathrm{SR_{even}}$. Parametrically bootstrapped 95\% confidence intervals for the slope did not overlap zero (Fig.~\ref{fig:SRmultReg}). The slope was weakly positive or not significantly different from zero in the most shade-adapted life forms (geophytes and phanerophytes), albeit the patterns were distinct in these groups. There were both hypostomatous ($\mathrm{SR_{even}} \approx 0$) and amphistomatous ($\mathrm{SR_{even}} \approx 1$) geophytes, but these were distributed across \el s. In contrast, phanerophytes were nearly always hypostomatous regardless of \el. % Allowing slopes to vary across life form signicantly increased model fit (lower AIC, Table~\ref{table:srAIC}).

@@

These observations from the British flora partially support the hypothesis that amphistomy can be used a proxy for open habitat in paleoenvironment reconstruction \citep{Carpenter_1994, Jordan_etal_2014, Carpenter_etal_2015} but also point out previously unknown subtleties. These previous studies based their conclusions on data from Proteaceae, in which there is little quantitative variation in stomatal ratio; species are either completely hypostomatous ($\mathrm{SR_{propAd}} \approx 0$) or completely amphistomatous ($\mathrm{SR_{propAd}} \approx 0.5$). Stomatal ratio in British angiosperms is also bimodal \citep{Peat_Fitter_1994b}, but across many families there is also quantitative variation. Importantly, this means that quantitative variation in stomatal ratio may provide a more precise, quantitative indicator of vegetation type, rather than simply `open' or `closed'. A quantitative relationship between \el~and stomatal ratio has already been shown for herbaceous perennials \citep{Bucher_etal_2017}, but we now know that it holds among annuals (therophytes), subshrubs (chamaephytes), and, to a lesser extent, geophytes as well (Fig.~\ref{fig:SRmultReg}).

@@

\begin{figure}[ht]
\centerline{\includegraphics[width=0.5\textwidth]{figures/figure_SRmultReg.pdf}}
%\centerline{\includegraphics[angle=270]{figures/figure_SRmultReg.eps}} % Needed this to upload file to journal website
\caption{The effect of light on stomatal ratio depends on Raunki\ae r life form. Greater Ellenberg light indicator values (\el) are associated with greater stomatal ratio ($\mathrm{SR_{even}}$) in therophytes, hemicryptophytes, and chamaephytes but not geophytes and phanerophytes. The maximum likelihood slope from phylogenetic regression is given with statistical significance based on 10$^4$ parametric bootstrap samples. Numbers in parentheses next to Raunki\ae r life form are the sample sizes in the final dataset. Estimated slopes (solid line) and 95\% bootstrapped confidence intervals (gray polygon between dashed lines) are plotted against raw data. Points have been jittered for visual clarity.} 
\label{fig:SRmultReg}
\end{figure}

@@

\end{document}

>>>>

\subsection*{Interactions between light and growth form determine stomatal ratio}

Overall, $\mathrm{SR_{even}}$ increased with \el, but there was a significant interaction ($\Delta\mathrm{AIC} > 2$, Table~\ref{table:srAIC}) between Raunki\ae r life form and \el~(Fig.~\ref{fig:SR-lf}). When classified as annual, biennial, perennial, shrub, or tree, growth form interacted with \el less ($\Delta \textrm{AIC} = \Sexpr{round(AIC(fitSR_gf_g) - AIC(fitSR_gf_f), 1)}$; Fig.~\ref{fig:SR-gf}). Raunki\ae r life form explained variation in stomatal ratio best (lower AIC; Table~\ref{table:srAIC}) , therefore we focus hereafter on those analyses. Both life form and \el~significantly increased model fit, though \el~had a markedly larger effect on model AIC (Table~\ref{table:srAIC}). The significant interaction is caused by different slopes between life forms. Among life forms with the overall greatest \el~(therophytes, hemicryptophytes, and chamaephytes, see Fig.~\ref{fig:lf-light}), there was a strong positive relationship between \el~and $\mathrm{SR_{even}}$. Parametrically bootstrapped 95\% confidence intervals for the slope did not overlap zero (Fig.~\ref{fig:SR-lf}). The slope was weakly positive or not significantly different from zero in the most shade-adapted life forms (geophytes and phanerophytes), albeit the patterns were distinct in these groups. There were both hypostomatous ($\mathrm{SR_{even}} \approx 0$) and amphistomatous ($\mathrm{SR_{even}} \approx 1$) geophytes, but these were distributed across \el s. In contrast, phanerophytes were nearly always hypostomatous regardless of \el. % Allowing slopes to vary across life form signicantly increased model fit (lower AIC, Table~\ref{table:srAIC}).

@@

These observations from the British flora partially support the hypothesis that amphistomy can be used a proxy for open habitat in paleoenvironment reconstruction \citep{Carpenter_1994, Jordan_etal_2014, Carpenter_etal_2015} but also point out previously unknown subtleties. These previous studies based their conclusions on data from Proteaceae, in which there is little quantitative variation in stomatal ratio; species are either completely hypostomatous ($\mathrm{SR_{propAd}} \approx 0$) or completely amphistomatous ($\mathrm{SR_{propAd}} \approx 0.5$). Stomatal ratio in British angiosperms is also bimodal \citep{Peat_Fitter_1994b}, but across many families there is also quantitative variation. Importantly, this means that quantitative variation in stomatal ratio may provide a more precise, quantitative indicator of vegetation type, rather than simply `open' or `closed'. A quantitative relationship between \el~and stomatal ratio has already been shown for herbaceous perennials \citep{Bucher_etal_2017}, but we now know that it holds among annuals (therophytes), subshrubs (chamaephytes), and, to a lesser extent, geophytes as well (Fig.~\ref{fig:SR-lf}).

@@

\begin{figure}[ht]
\centerline{\includegraphics[width=0.5\textwidth]{figures/figure_SR-lf.pdf}}
%\centerline{\includegraphics[angle=270]{figures/figure_SRmultReg.eps}} % Needed this to upload file to journal website
\caption{The effect of light on stomatal ratio depends on Raunki\ae r life form. Greater Ellenberg light indicator values (\el) are associated with greater stomatal ratio ($\mathrm{SR_{even}}$) in therophytes, hemicryptophytes, and chamaephytes but not geophytes and phanerophytes. The maximum likelihood slope from phylogenetic regression is given with statistical significance based on 10$^4$ parametric bootstrap samples. Numbers in parentheses next to Raunki\ae r life form are the sample sizes in the final dataset. Estimated slopes (solid line) and 95\% bootstrapped confidence intervals (gray polygon between dashed lines) are plotted against raw data. Points have been jittered for visual clarity.} 
\label{fig:SR-lf}
\end{figure}

@@

\begin{figure}[ht]
\centerline{\includegraphics[width=0.5\textwidth]{figures/figure_SR-gf.pdf}}
%\centerline{\includegraphics[angle=270]{figures/figure_SRmultReg.eps}} % Needed this to upload file to journal website
\caption{The effect of light on stomatal ratio depends on growth form. Greater Ellenberg light indicator values (\el) are associated with greater stomatal ratio ($\mathrm{SR_{even}}$) in annual, biennual, and perennial herbs, but not shrubs or trees. The maximum likelihood slope from phylogenetic regression is given with statistical significance based on 10$^4$ parametric bootstrap samples. Numbers in parentheses next to growth form are the sample sizes in the final dataset. Estimated slopes (solid line) and 95\% bootstrapped confidence intervals (gray polygon between dashed lines) are plotted against raw data. Points have been jittered for visual clarity.} 
\label{fig:SR-gf}
\end{figure}
\end{document}

### 2017-10-05 ms/ms.tex
updated results

Ellenberg light indicator values (\el) differed significantly among growth forms. Among Raunki\ae r life forms, therophytes (annuals), hemicryptophytes (perennial herbs with buds near the soil surface), and chamaephytes (subshrubs) had greater \els~than phanerophytes (woody plants) and geophytes (perennial herbs with storage organs) (Fig.~\ref{fig:lf-light}; ANOVA - $F_{4, 367}$ = 18.3, $P$ = 1.1 $\times10^{-13}$). Likewise, herbaceous plants (annual, biennial, and perennials) had greater \els~ than shrubs and trees (Fig.~\ref{fig:gf-light}; ANOVA - $F_{4, 367}$ = 10.8, $P$ = 2.6 $\times10^{-8}$)

\subsection*{Interactions between light and Raunki\ae r life form determine stomatal ratio}

Overall, $\mathrm{SR_{even}}$ increased with \el, but there was a significant interaction ($\Delta\mathrm{AIC} > 2$, Table~\ref{table:srAIC}) between Raunki\ae r life form and \el~(Fig.~\ref{fig:SRmultReg}). Both life form and \el~significantly increased model fit, though \el~had a markedly larger effect on model AIC (Table~\ref{table:srAIC}). The significant interaction is caused by different slopes between life forms. Among life forms with the overall greatest \el~(therophytes, hemicryptophytes, and chamaephytes, see Fig.~\ref{fig:lf-light}), there was a strong positive relationship between \el~and $\mathrm{SR_{even}}$. Parametrically bootstrapped 95\% confidence intervals for the slope did not overlap zero (Fig.~\ref{fig:SRmultReg}). The slope was weakly positive or not significantly different from zero in the most shade-adapted life forms (geophytes and phanerophytes), albeit the patterns were distinct in these groups. There were both hypostomatous ($\mathrm{SR_{even}} \approx 0$) and amphistomatous ($\mathrm{SR_{even}} \approx 1$) geophytes, but these were distributed across \el s. In contrast, phanerophytes were nearly always hypostomatous regardless of \el. % Allowing slopes to vary across life form signicantly increased model fit (lower AIC, Table~\ref{table:srAIC}).

>>>>

Ellenberg light indicator values (\el) differed significantly among growth forms. Among Raunki\ae r life forms, therophytes (annuals), hemicryptophytes (perennial herbs with buds near the soil surface), and chamaephytes (subshrubs) had greater \els~than phanerophytes (woody plants) and geophytes (perennial herbs with storage organs) (Fig.~\ref{fig:lf-light}; ANOVA - $F_{4, 367}$ = 18.3, $P$ = 1.1 $\times10^{-13}$). Likewise, herbaceous plants (annual, biennial, and perennials) had greater \els~than shrubs and trees (Fig.~\ref{fig:gf-light}; ANOVA - $F_{4, 367}$ = 10.8, $P$ = 2.6 $\times10^{-8}$)

\subsection*{Interactions between light and growth form determine stomatal ratio}

Overall, $\mathrm{SR_{even}}$ increased with \el, but there was a significant interaction ($\Delta\mathrm{AIC} > 2$, Table~\ref{table:srAIC}) between Raunki\ae r life form and \el~(Fig.~\ref{fig:SR-lf}). When classified as annual, biennial, perennial, shrub, or tree, growth form interacted with \el less ($\Delta \textrm{AIC} = 2.4$; Fig.~\ref{fig:SR-gf}). Raunki\ae r life form explained variation in stomatal ratio best (lower AIC; Table~\ref{table:srAIC}) , therefore we focus hereafter on those analyses. Both life form and \el~significantly increased model fit, though \el~had a markedly larger effect on model AIC (Table~\ref{table:srAIC}). The significant interaction is caused by different slopes between life forms. Among life forms with the overall greatest \el~(therophytes, hemicryptophytes, and chamaephytes, see Fig.~\ref{fig:lf-light}), there was a strong positive relationship between \el~and $\mathrm{SR_{even}}$. Parametrically bootstrapped 95\% confidence intervals for the slope did not overlap zero (Fig.~\ref{fig:SR-lf}). The slope was weakly positive or not significantly different from zero in the most shade-adapted life forms (geophytes and phanerophytes), albeit the patterns were distinct in these groups. There were both hypostomatous ($\mathrm{SR_{even}} \approx 0$) and amphistomatous ($\mathrm{SR_{even}} \approx 1$) geophytes, but these were distributed across \el s. In contrast, phanerophytes were nearly always hypostomatous regardless of \el. % Allowing slopes to vary across life form signicantly increased model fit (lower AIC, Table~\ref{table:srAIC}).
